Convergent Adaptation while adapting

### 2014-06-30 patchy-selection-paper.tex
put in dispersal refs

@@

The convergent evolution of similar phenotypes in response to shared
selection pressures is a testiment to the power of selection to
repeatedly sculpt phenotypic variation. 
In some cases this convergence extends to the molecular level, 
with disparate taxa converging to the same phenotype through parallel 
genetic changes in the same pathway, genes, 
or even by precisely the same genetic changes \citep{zhen2012parallel}. 
Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
sometimes over deep time scales \citep{deephomologypapers}. 
Convergence on a molecular level also suggests that the path 
accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
as would be the case if only a few changes confer the advantagous phenotype 
and are sufficiently free of deleterious pleiotropic consequences.

%For highly polygenic traits

Convergent adaptation can also occur within species, if individuals within a species 
adapt to the same environment, in parallel, through different genotypic changes. 
There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
Such evolution of convergent phenotypes is favored by higher mutational input, 
i.e.\ larger mutational target and/or population size \citep{dmitri}.
The geographic distribution of populations can also affect the probability of parallel mutation within a species:
a widespread species is more likely to adapt by multiple, parallel mutations if dispersal is geographically limited, 
since subpopulations will adapt via new mutations before the adaptive allele arrives via migration \citep{ralph2010parallel}. 
Standing variation for a trait can also increase the probability of convergence, 
and standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
(beginning from a base of multiple copies),
leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.

@@

One elegant set of such examples is provided by the assortment of plant species 
that have repeatedly adapted to patches of soil with high concentrations of heavy metals
(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine,mimulus};
the alleles conferring heavy metal tolerance are thought to be locally restricted 
because they incur a cost off of these patches. 
Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson1933concealing}.
One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice1940ecologic}.
Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
and, perhaps as a result of this strong selection against migrants, 
at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman2003different}. 
Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
and in lizards \citep{whitesands}.

In this paper, we study this situation, namely,
when a set of alleles provide an adaptive benefit in geographically localized patches, 
separated by (inhabited) regions where the alleles are deleterious.
\plr{or neutral?}
The main questions are:
Under what conditions is it likely that each patch adapts in parallel, through new mutation,
and when is it likely that migration carries these alleles between patches?
How could we tell, after the fact, which has occurred?
We work in a model of continuous geography,
using a variety of known results and new methods.
In section XX we \dots QUICK OUTLINE

@@

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in spatially varying selection models. 
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{s_m}$. Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panimictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}.

@@

Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(supposing that no $B$ allele arises independently by mutation in the meantime).
To determine this, we dissect the process by which an allele transits between the patches,
in the process obtaining other useful information about the genealogy of B alleles.

@@

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
is $\gamma(x)$.
To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
of the outflux $\gamma$ at that point, mulitplied by the relevant second quantity.
By examining the constituent terms,
it will turn out that equation \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q}.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
gives the expected total number of individuals ever living in $A$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
\end{align}
Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
which we assume is circular, has relatively small area $A$ and is located near position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-\alpha t}/\E[K]$.
It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
as in \citet{barton1987establishment};
but we make a simpler approximation, approximating this probability by the chance that the ``trunk'' hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity,
successfulling establishing in the new patch.
Write $h(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h(x,y)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
\end{align}

Intuitively, $h(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see Appendix~\ref{apx:hitting_occupation}) that
\begin{align} 
  h(x,y) = \frac{ g(A) }{ \E[K] } u(x,y),
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives,
and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
(at least asymptotically as $x$ and $y$ are far apart),
we now know that the outflux of successful migrant families is equal to the mean occupation frequency
multiplied by $g(A) f(A) / \E[K]$.
The probability $f(A)$ that the family does not die out once it arrives at the new patch is
$\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
if each member arriving has chance $p_e$ of establishing, independently.
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
% \end{align}
Combining the terms with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
we obtain that
\begin{align}
  \label{eqn:migrate} 
  \migrate(y) &\approx q(y) \times \frac{ \rho s_p  }{ \xi^2 s_m }  \\
  % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) .
\end{align}
This does not depend on $A$, because although
$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak:
for instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, and correction terms, see Appendix~\ref{apx:hitting_occupation}.

@@

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and benefical selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
\begin{equation} \label{eqn:parallel_prob}
  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A
    \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.

%%%%%%%%%%%%
\subsection{Multiple patches}
Above we worked with two patches separated by distance $R$;
but these calculations apply as well to multiple patches
in which the $B$ allele can facilitate local adaptation. 
Under our assumptions of mutational exclusion and rapid establishment, 
each patch will adapt through establishment of a single copy of the $B$ allele, 
either by migration or mutation.
If we imagine that the times between successive establishments is the result of many nearly-independent attempts
with small probability of success,
the process of occupation is well approximated by a Markov chain,
whose state records the labeling of patches according to the colonizing allele
(or none, for those that haven't yet adapted).

@@

When an already adapted patch colonizes another through migration,
then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
The genetic length of this haplotype is roughly inversely proportional to the
number of generations the allele spends ``in transit'', 
because each reproductive event in locations where the allele is at
low frequency will likely recombine haplotypes of the $b$ allele closer to the adaptive allele. 
In this case, the haplotype is literally hitchhiking across geography!
Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
including the lineage that founds an adaptive population on a second patch,
and the length of the haplotype shared between this lineage and one in the original patch.
The entire time that the lineage is outside the region where the B allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
the haplotype that accompanies it is broken down rapidly. 
After the lineage establishes on the patch, the rate of decay of the haplotype is slowed significantly, 
since most others with which it recombines have similar haplotypic backgrounds. 
We first discuss the transit of the lineage, and its accompanying haplotype, and return
to how haplotypes are whittled down within estalbished patches in the next section.

Above we argued that a good model for this transit is a single Brownian trunk lineage
surrounded by a cloud of close relatives of random size $K$
whose chance of surviving until $t$ generations is $k_e(t)$.
Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
with $\tau = \infty$ if it never does. 
We assume that the length of the hitchhiking segment depends only on 
the time spent by the trunk lineage of the first successful migrant ``in transit'' between the patches,
i.e.\ $\tau$ conditioned on $\tau < \infty$.
In so doing, we neglect 
recombination between B alleles (since they are at low density in transit, we assume this is rare),
and the possibility that more than one migrant family is in transit at once
(so that faster migrants would be more likely to have arrived first);
but these factors should be minor if the rest of our framework fits.

Our first obstacle is that we do not know $k_e(t)$, 
the distribution of the time until a family dies out, exactly.
However, we do know that $k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
replacing this random time by an exponential with rate $s_m$.
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
In particular, if the patch is circular with radius $w$ and a distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

Refer again to Figure~\ref{fig:lineagesmotion}. 
When a lineage is in a region with high frequency of B alleles (dark red), 
it is very unlikely to recombine off the B background. 
However, when it wanders out of the patch,
into regions with lower frequency of B alleles,
when it recombines it will
usually be on to the background of the b allele.

>>>>

\title{
Convergent Adaptation while adapting 
}

\author{Peter L. Ralph$^1$ \\ email: pralph@usc.edu  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }

@@

% Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
% sometimes over deep time scales \citep{deephomologypapers}. 
% Convergence on a molecular level also suggests that the path 
% accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
% as would be the case if only a few changes confer the advantagous phenotype 
% and are sufficiently free of deleterious pleiotropic consequences.

The convergent evolution of similar phenotypes in response to shared
selection pressures is a testiment to the power of selection to
repeatedly sculpt phenotypic variation. 
In some cases this convergence extends to the molecular level, 
with disparate taxa converging to the same phenotype through parallel 
genetic changes in the same pathway, genes, 
or even by precisely the same genetic changes \citep{zhen2012parallel,,martin2013repeated,deephomologypapers}. 
Convergent adaptation can also occur within species, if individuals within a species 
adapt to the same environment, in parallel, through different genetics changes. 
There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
Such evolution of convergent phenotypes is favored by higher mutational input, 
i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
The geographic distribution of populations can also affect the probability of parallel mutation within a species:
a widespread species is more likely to adapt by multiple, parallel mutations if dispersal is geographically limited, 
since subpopulations will adapt via new mutations before the adaptive allele arrives via migration \citep{ralph2010parallel}. 
Standing variation for a trait can also increase the probability of convergence, 
and standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
(beginning from a base of multiple copies),
leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.

@@

One elegant set of such examples is provided by the assortment of plant species 
that have repeatedly adapted to patches of soil with high concentrations of heavy metals
(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine};
the alleles conferring heavy metal tolerance are thought to be locally restricted 
because they incur a cost off of these patches \citep{wright2013indirect}.
Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson1933concealing}.
\plr{we are missing what Nachman said was the original ref for this... I think you took a picture?}
One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice1940ecologic}.
Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
and, perhaps as a result of this strong selection against migrants, 
at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
and in lizards \citep{rosenblum2010molecular}.

In this paper, we study this situation, namely,
when a set of alleles provide an adaptive benefit in geographically localized patches, 
separated by (inhabited) regions where the alleles are deleterious.
The main questions are:
Under what conditions is it likely that each patch adapts in parallel, through new mutation,
and when is it likely that migration carries these alleles between patches?

How could we tell, after the fact, which has occurred?

We work in a model of continuous geography,
using a variety of known results and new methods.
In section \ref{ss:patchymutation} we develop a simple approximation,
eqn. \eqref{eqn:mutrate}, for the rate at which patches become
established by new mutations. Then in section \ref{ss:patchymigration} we
develop an approximation, eqn. \eqref{eqn:migrate}, to the rate which alleles move from an
existing patch to colonize a novel patch (this represents a big chunk
of the results of the paper). We combine these two results in section
\ref{ss:probparallel} to develop an approximation to the probability of
parallel adaptation, eqn. \eqref{}.
To understand the genomic signal left by adaptations shared by
migration in section \ref{ss:haplotype_length} we approximate the time
it will take for an allele to transit between patches,
eqn. \eqref{eqn:mean_tau}, and use this to approximate the length of
haplotype that hitchhikes with it,
eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
haplotype decays within patches eqn. \eqref{} in section \ref{ss:patch_haplotype}.

\dots QUICK OUTLINE

@@

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in spatially varying selection models. 
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{s_m}$. Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panimictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
More general work on the mathematically equivalent problem of density-dependent population growth 
has obtained much more general criteria
under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.

@@

Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
we would like to know the time scale on which copies of $B$ move between patches by migration
(supposing that no $B$ allele arises independently by mutation in the meantime).
To determine this, we dissect the process by which an allele transits between the patches,
in the process obtaining other useful information about the genealogy of $B$ alleles.

@@

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
is $\gamma(x)$.
To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
of the outflux $\gamma$ at that point, multiplied by the relevant second quantity.
By examining the constituent terms,
it will turn out that equation \eqref{eqn:gestalt_migrate} 
is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
Concretely, it is the density of $\E[\int_0^t Z_t(\cdot) dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set $S$, 
gives the expected total number of individuals ever living in $S$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
\end{align}
% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
which we assume is circular, has relatively small area $A$ and is centered at position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
It would be possible to analyze directly the probability that a given migrant family establishes in the patch,
as in \citet{barton1987establishment};
but we make a simpler approximation, approximating this probability by the chance that the ``trunk'' hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity,
successfully establishing in the new patch.
Write $h_A(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h_A(x,y)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
\end{align}

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see Appendix~\ref{apx:hitting_occupation}) that
\begin{align} 
  h_A(x,y) \approx \frac{ g(A) }{ \E[K] } u(x,y),
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives,
and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
The probability $f(A)$ that the family does not die out once it arrives at the new patch is
$\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
if each member arriving has chance $p_e$ of establishing, independently.
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
% \end{align}
Combining with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
we obtain that
\begin{align}
  \label{eqn:migrate} 
  \migrate(y) &\approx q(y) \times \frac{ \rho s_p  }{ \xi^2 s_m }  \\
  % \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) .
\end{align}
This does not depend on $A$, because although
$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak:
for instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see Appendix~\ref{apx:hitting_occupation}.

@@

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and benefical selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
\begin{equation} \label{eqn:parallel_prob}
  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu s_m }{2 A \mu s_m + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.

%%%%%%%%%%%%
\subsubsection{Multiple patches}
Above we worked with two patches separated by distance $R$;
but these calculations apply as well to multiple patches
in which the $B$ allele can facilitate local adaptation. 
Under our assumptions of mutational exclusion and rapid establishment, 
each patch will adapt through establishment of a single copy of the $B$ allele, 
either by migration or mutation.
If we imagine that the times between successive establishments is the result of many nearly-independent attempts
with small probability of success,
the process of occupation is well approximated by a Markov chain,
whose state records the labeling of patches according to the colonizing allele
(or none, for those that haven't yet adapted).

@@

When an already adapted patch colonizes another through migration,
then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
The genetic length of this haplotype is roughly inversely proportional to the
number of generations the allele spends ``in transit'', 
because each reproductive event in locations where the allele is at
low frequency will likely recombine haplotypes of the $b$ allele closer to the adaptive allele. 
In this case, the haplotype is literally hitchhiking across geography!
Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
including the lineage that founds an adaptive population on a second patch,
and the length of the haplotype shared between this lineage and one in the original patch.
The entire time that the lineage is outside the region where the $B$ allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
the haplotype that accompanies it is broken down rapidly. 
After the lineage establishes on the patch, the rate of decay of the haplotype is slowed significantly, 
since most others with which it recombines have similar haplotypic backgrounds. 
We first discuss the transit of the lineage, and its accompanying haplotype, and return
to how haplotypes are whittled down within estalbished patches in the next section.

Above we argued that a good model for this transit is a single Brownian trunk lineage
surrounded by a cloud of close relatives of random size $K$
whose chance of surviving until $t$ generations is $1-k_e(t)$.
Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
with $\tau = \infty$ if it never does. 
We assume that the length of the hitchhiking segment depends only on 
the time spent by the trunk lineage of the first successful migrant ``in transit'' between the patches,
i.e.\ $\tau$ conditioned on $\tau < \infty$.
In so doing, we neglect 
recombination between $B$ alleles (since they are at low density in transit, we assume this is rare),
and the possibility that more than one migrant family is in transit at once
(so that faster migrants would be more likely to have arrived first);
but these factors should be minor if the rest of our framework fits.

Our first obstacle is that we do not know $1-k_e(t)$, 
the probability a family survives until $t$, exactly.
However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
replacing this random time by an exponential with rate $s_m$.
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
In particular, if the patch is circular with radius $w$ and a distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

Refer again to Figure~\ref{fig:lineagesmotion}. 
When a lineage is in a region with high frequency of $B$ alleles (dark red), 
it is very unlikely to recombine off the $B$ background. 
However, when it wanders out of the patch,
into regions with lower frequency of $B$ alleles,
when it recombines it will
usually be on to the background of the $b$ allele.
